---
title: "Fatal Police Shootings"
output:
  pdf_document: default
  html_document:
    code_folding: hide
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r,messages=FALSE,results='hide',warnings=FALSE,include=FALSE}
###Data Structure & Summary
options(warn=-1,verbose=FALSE)
library(ggplot2)
library(gridExtra)
library(maps)
library(data.table)
library(lubridate)
library(RColorBrewer)
library(corrplot)

path='/Users/Mikey/Documents/ML-Case-Studies/US Police Shooting Fatalities'
setwd(path)

filenames<-c('Guardian_2015.csv','Guardian_2016.csv','WP_database.csv')
guard2015.dt<-fread(filenames[1],sep=',')
guard2016.dt<-fread(filenames[2],sep=',')

```


```{r,include=FALSE}
##Data Cleaning & Miscellaneous Functions

clean.Guard2015 <-function(datafile){
    datafile[,age:=ifelse(age=='Unknown',-1,age)] #-1: Unknown
    
    datafile[,uid:=as.numeric(uid)]
    datafile[,age:=as.numeric(age)]
    datafile[,gender:=as.factor(gender)]
    
    datafile[,race:=as.factor(raceethnicity)]
    datafile[,raceethnicity:=NULL]
    
    datafile[,armed:=as.factor(armed)]
    datafile[,day:=as.factor(day)]
    datafile[,state:=as.factor(state)]
    datafile[,classification:=as.factor(classification)]
    
    datafile
}

clean.Guard2016<-function(datafile){
    datafile[,age:=ifelse(age=='Unknown',-1,age)] #-1: Unknown
    datafile[,age:=ifelse(age=='40s',-2,age)] #-1: 'Unknown'40s'

    datafile[,uid:=as.numeric(uid)]
    datafile[,age:=as.numeric(age)]
    datafile[,gender:=as.factor(gender)]

    datafile[,race:=as.factor(raceethnicity)]
    datafile[,raceethnicity:=NULL]
    
    datafile[,armed:=as.factor(armed)]
    datafile[,day:=as.factor(day)]
    datafile[,state:=as.factor(state)]
    datafile[,classification:=as.factor(classification)]
    
    datafile
}


tbl.ECDF<-function(datatable){
    ECDF<-vector(mode='numeric',length = length(datatable))
    
    for (i in 1:nrow(datatable)) {
       ECDF[i]<-sum(datatable[nrow(datatable):(nrow(datatable)-(i-1))])    
    }
    
    ECDF<-ECDF/sum(datatable)
}

```


```{r,include=FALSE}
#Create Single data table for each source
clean.guard.dt<-rbindlist(
                    list(
                        clean.Guard2015(guard2015.dt),
                        clean.Guard2016(guard2016.dt)
                        ),
                    use.names=TRUE
                    )
    clean.guard.dt[,year:=as.factor(year)]

```

##Exploratory Data Analysis

The datacan be broadly classified into two groups: victim descriptors and situation sescriptors. Victim descriptors provide information about the victim that is independent of the specific altercation that resulted in death. Situation descriptors describe the immediate environment of the incident and the interaction between the involved parties, as described by the 'Armed' and 'Classification' fields.

Splitting the descriptors into these groups is useful because the victim and situation information answer the questions "What people are most frequently victims of shootings?" and "What situations result in police gun shooting fatalities?", respectively.


*Victim Descriptors*    |*Situation Descriptors*
------------            |------------- 
Name                    |Date
Age                     |Location (State & City)
Race                    |Classification (Manner of Death)
Gender                  |Armed (Type of Weapon on Victim)

###Victim Data

The age distribution of police-related fatalities is nearly identical for both 2015 and 2016 with a median value of 35 years. The tail at the higher ages indicates that the chance of police altercations becoming lethal reduces gradually as people become older. This may be due to a decreased inclination for police to use lethal force or a decreased ability for the person to present himself as a significant threat. 

```{r}
##Victim

    #Age Distribution
    g.guard.age<-ggplot(data = clean.guard.dt[age>=0],aes(x=age,fill=year))+
        geom_histogram(position='identity',alpha=0.5,binwidth = 5)+
        xlab('Victim Age')+ylab('Fatalities')+theme(text = element_text(size = 15))+
        annotate("text", x = 70, y = 150, 
                 label = paste('Median Age = ',median(clean.guard.dt[age>=0]$age)))
    
    g.guard.age
```

The race of an individual appears to be a significant indicator of police-related fatalities. There is a clear disparity with 3 of the 8 reported races accounting for 94.2% of all reported fatalities. It should be noted that the relative values of the reported need to be adjusted for population before the effects of race on odds of police-related fatality can be estimated.
```{r,fig.height=5,fig.width=8}
    #Race Distribution
    g.guard.race<-ggplot(data=clean.guard.dt, aes(x=reorder(race, table(race)[race]))) + 
        geom_bar(fill='red')+
        xlab('Race')+ylab('Total Fatalities 2015-2016')+
        theme(text = element_text(size = 15))+
        coord_flip()

    g.guard.race

    #Race ECDF
    guard.tbl.race<- table(clean.guard.dt$race)
    guard.tbl.race<-guard.tbl.race[order(guard.tbl.race)]
    guard.race.ECDF<-tbl.ECDF(guard.tbl.race) #ECDF of WP weapons
    
    print(paste0('White, Black, & Hispanic Deaths Account for ', 
                        as.character(round(guard.race.ECDF[3]*100,digits=1)),
                            '% of all fatalities' ))
```

Of the 3 victim descriptors, gender is the most significant predictor with men being 20 times more likely of being killed by police than women. 
```{r}
    clean.guard.dt[,.N,by=gender]
```

###Situation Data

The distribution of fatalities across the United States indicates that there is a much higher occurrence of police-related deaths in California, Texas, Florida, and Arizona compared to the rest of the states. If the fatalities in these top 4 states are ignored, the number of deaths decreases slowly from across states implying that police-related deaths occur at the same rate per year regardless of location.
```{r,fig.height=10,fig.width=12}
##Situation
    #Location (States)
    g.guard.state<-ggplot(data=clean.guard.dt, 
                          aes(x=reorder(state, table(state)[state]))) + 
        geom_bar(fill='red')+coord_flip()+
        xlab('State')+ylab('Fatalities')+
        theme(text = element_text(size = 15))
    
    g.guard.state    
```

A plot of the type of weapon in possession by victims is intriguing. It makes sense that the weapon that is most frequently associated with police using lethal force is a firearm. After firearms however, most fatalities occur when people are unarmed. This is important because it means that even when a person does not have a weapon, there exist conditions when police will deem it necessary to use lethal force. 
```{r,fig.width=12}
    #Armed
    g.guard.armed<-ggplot(data=clean.guard.dt, 
                          aes(x=reorder(armed, table(armed)[armed]),fill=race)) + 
        geom_bar()+ scale_fill_brewer(palette = 'BrBG')+
        coord_flip()+
        xlab('Type of Weapon')+ylab('Fatalities')+
        theme(text = element_text(size = 15))
    
   
 
    g.guard.armed 
```


The method used by police that most frequently results in a fatality is gunshot. Of the 2226 documented fatalities, 2021 were the result of firearm use by the police. All analyses should be done in the context of gunshot fatalities because those deaths essentially describe the entire dataset with the other classifications providing relatively little information.
```{r,fig.width=12}
 #Cause of Death
    g.guard.COD<-ggplot(data=clean.guard.dt, 
                          aes(x=reorder(classification, 
                                        table(classification)[classification]),fill=race)) + 
        geom_bar()+coord_flip()+
        xlab('Cause of Death')+ylab('Fatalities')+
        scale_fill_brewer(palette = 'BrBG')+
        theme(text = element_text(size = 15))

    g.guard.COD     
```


```{r,eval=FALSE,include=FALSE}    
#Top 95% of Weapons in Possession of Killed Citizen
guard.tbl_armed<- table(clean.guard.dt$armed)
    guard.tbl.armed<-guard.tbl_armed[order(guard.tbl.armed)]
    guard.top10.armed<-tail(guard.tbl.armed,n=6)
    
    print(
      paste(
              'Top 10 Types of Weapons Account for' ,
                    as.character(sum(guard.top10.armed)/sum(guard.tbl.armed)*100) ,
                    '% of National Occurrences'
            ),sep = ''
      )
  

guard.tbl.state<-table(clean.guard.dt[classification=='Gunshot']$state)
      guard.tbl.state<-guard.tbl.state[order(guard.tbl.state)]
      N=10
      guard.top.states<-tail(guard.tbl.state,n=N)
    
      print(paste('Median 2015-2016 State Deaths by Police Shootings is ',
                  median(guard.tbl.state) ))
    
      print(
       paste(
              as.character(N), 'States with the Most Police Shootings Account for' ,
                    as.character(sum(guard.top.states)/sum(guard.tbl.state)*100) ,
                    '% of Police Shooting Fatalities'
            ),sep = ''
      )

```

###Analysis of Descriptor Interactions

Direct analysis of the victim and situation descriptors provides a coarse view of the possible predictors of police-related fatalities. They are not useful in isolation because when a police officer decides to use lethal force, he must consider multiple conditions that are occurring simultaneously. An example would be 'Is my attacker physically strong and does he possess a weapon?'.

Since most of the variables in the dataset are not numeric, a standard correlation plot cannot be used to identify multiple conditions that describe a single fatality. However, it can be analyzed using a heatmap that indicates how many times pairs of conditions occur. The presented heatmap indicates this for male gunshot victims. Victim age has been binned into 5 year increments. Red and yellow regions indicate low and high occurrence of the condition pairs, respectively. 

**Variables of Interest:** Age, Month, Type of Weapon, & Race

```{r,fig.height=10,fig.width=15}

cond.target<-function(field, target){
    #(target-target%%5)/5 + 1
    
    age.levels <- as.character(c(1:17))
    
    month.levels <- c(" 'January' " ," 'February' ", " 'March' "," 'April' "," 'May' ", " 'June' ", " 'July' ",
                        " 'August' "," 'September' ", " 'October' ", " 'November' ", " 'December' ")
    
    armed.levels<-c( " 'Disputed' " ," 'Firearm' "," 'Knife' "," 'No' ",
                     " 'Non-lethal firearm' "," 'Other' "," 'Unknown' "," 'Vehicle' ")           

    race.levels<-c(" 'Arab-American' "," 'Asian/Pacific Islander' ", " 'Black' "," 'Hispanic/Latino' ",       
                    " 'Native American' "," 'Other' "," 'Unknown' "," 'White' " )
    
    v.str <- ifelse(field==1,'((age - age%%5)/5)',
                  ifelse(field==2,'month',
                         ifelse(field==3,'armed','race')))
    
    trgt.str<-ifelse(field==1,age.levels[target],
                  ifelse(field==2,month.levels[target],
                         ifelse(field==3,armed.levels[target],race.levels[target])))
    
    paste0(v.str,'==',trgt.str)
} #Creates Conditional Indexing String


#v('gunshot' & 'male') = [ [age], [month] ,[armed], [race] ]
#age.levels <- as.character(c(1:17))
age.levels <- c('Age:5-10','Age:11-15','Age:16-20',
                'Age:21-25','Age:26-30','Age:31-35','Age:36-40',
                'Age:41-45','Age:46-50','Age:51-55','Age:56-60',
                'Age:61-65','Age:66-70','Age:71-75','Age:76-80',
                'Age:81-85','Age:86-90')

month.levels <- c( 'Month:January'  , 'Month:February' ,  'Month:March' , 'Month:April' , 'Month:May' , 
                   'Month:June' ,  'Month:July' ,  'Month:August' , 'Month:September' , 
                   'Month:October' ,  'Month:November' , 'Month:December' )

armed.levels<-c(  'Armed:Disputed'  , 'Armed:Firearm' , 'Armed:Knife' , 'Armed:No' ,
                  'Armed:Non-lethal firearm' , 'Armed:Other' , 'Armed:Unknown' , 'Armed:Vehicle' )           

race.levels<-c( 'Race:Arab-American' , 'Race:Asian/Pacific Islander' ,  'Race:Black' ,
                'Race:Hispanic/Latino' ,  'Race:Native American' ,
                'Race:Other' , 'Race:Unknown' , 'Race:White'  )

pred.names<-c(age.levels,month.levels,armed.levels,race.levels)

age.bins.N <-length(age.levels)
month.N <- length(month.levels)
armed.N <- length(armed.levels)
race.N<-length(race.levels)
total.levels<-age.bins.N+month.N+armed.N+race.N

mat.heat<-matrix(nrow=total.levels,ncol=total.levels)

for (i in 1:total.levels) {
    cond1.field<-ifelse(i<age.bins.N+1,1,
                    ifelse(i<age.bins.N+month.N+1,2,
                        ifelse(i<age.bins.N+month.N+armed.N+1,3,4)))
    
    cond1.target<-ifelse(i<age.bins.N+1,i,
                    ifelse(i<age.bins.N+month.N+1,i-age.bins.N,
                        ifelse(i<age.bins.N+month.N+armed.N+1,i-age.bins.N-month.N,
                                        i-age.bins.N-month.N-armed.N)))
    
    for (j in 1:total.levels) {
        cond2.field<-ifelse(j<age.bins.N+1,1,
                        ifelse(j<age.bins.N+month.N+1,2,
                            ifelse(j<age.bins.N+month.N+armed.N+1,3,4)))
            
        cond2.target<-ifelse(j<age.bins.N+1,j,
                    ifelse(j<age.bins.N+month.N+1,j-age.bins.N,
                        ifelse(j<age.bins.N+month.N+armed.N+1,j-age.bins.N-month.N,
                                        j-age.bins.N-month.N-armed.N)))
        
        cond1 <- parse(text=cond.target(cond1.field,cond1.target))
        cond2 <- parse(text=cond.target(cond2.field,cond2.target))
        
        if(j==i){#Artificially Zero Correlation with Self
            mat.heat[j,i]=0
            next
        }
        
        mat.heat[j,i]<- nrow(clean.guard.dt[((age>=0) & gender=='Male' & classification =='Gunshot' & 
                                                 eval(cond1) & eval(cond2))])   
    }    
}

heatmap(mat.heat+1,labCol=pred.names,labRow = pred.names,
        Rowv = NULL,Colv = NA,scale = 'column', 
        main='Male Gunshot Fatalities: Factor Interactions',
        margins=c(9,9),cexRow = 1.25,cexCol = 1.25)
```

####Top 50 Most Frequent Condition Pairs
```{r,results=FALSE,include=FALSE}
top50.vals<-tail(sort(mat.heat),n=100)
top50.vals<-top50.vals[seq(from=1,to=length(top50.vals),by=2)]

top.cond<-matrix(nrow=1,ncol=3)
for (i in 1:length(top50.vals)) {
    top.cond<- rbind(top.cond, c(which(mat.heat==top50.vals[i],arr.ind = TRUE)[1,],top50.vals[i]))
}

top.cond<-top.cond[2:nrow(top.cond),]
top.cond.dt<-data.table(top.cond)
setnames(top.cond.dt,c('C1','C2','Occurrences'))

top.cond.dt[,Condition1:=pred.names[C1]]
top.cond.dt[,Condition2:=pred.names[C2]]
setorderv(top.cond.dt,'Occurrences',-1)
```

```{r}
top.cond.dt[,.(Occurrences,Condition1,Condition2)]
```

